Mitigating Interest Flooding DDoS Attacks in Named Data Networking

### 2012-12-09 01-intro.tex
New intro, updated discussion
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@876 df1a90cd-247b-4364-bdaf-287a704eb499

Denial of service (DoS) and distributed denial of service attacks (DDoS) appeared not very long time after Internet got traction and since then a constant arms race is going on between information providers and attackers who want to knock a certain source of information down. Running a DDoS attack has become a very successful business model due to its general scalable nature, cheapness and lack of protection of information providers by government and law. Most popular and economically effective DDoS attacks are organized with botnets. A botnet agents get installed by many various ways on computers of ordinary people that leads to a very wide geographical distribution of the source of attack that makes blacklisting impossible (in contrast to single source DoS attacks).

In 2011 for the first time IPv6 DDoS attacks on commercial networks were observed ~\cite{Arbor}. They occur rarely, which only means that IPv6 is not widespread enough for attackers to put significant resources into such DDoS attacks. Principally, in terms of DDoS security IPv6 didn't advance that much comparing to standard IPv4 networks.

IP architecture has a wide surface for denial of service attacks which takes the beginning from the thin-waist IP protocol, goes through transport level and up to DNS level. In IP it is very easy to forge source address and specify the exact destination that needs to be placed under attack that creates possibilities for performing various types of reflection DoS attacks. On transport level in TCP/IP stack SYN flood, RST flood and FIN-way attacks are very effective and popular. DNS system can be attacked in multiple ways, e.g. DNS amplification attack.

We believe that future Internet architectures can do and must do more for DDoS mitigation. By its design NDN architecture is not sensitive to any kind of reflection attacks due to the lack of host addresses and ubiquitous data caching. NDN architecture can operate without any DNS-like architecture since it routes and forwards packets based on human-readable names instead of IP addresses that have been received from DNS.

%\todo{
%Current Internet architecture and its resilience to DoS and DDOS. Requirements from a new architecture such as CCN with respect to DoS and DDoS resilience.

%Our contributions.

%Paper organization.}

In this paper we provide a analyze all aspects of Interest Flooding attack in NDN architecture, propose multiple defensive techniques against such attacks, establish evaluation methodology using ndnSIM simulation software, execute small scale and Internet scale DDoS attack scenarios, and analyze our experimental results.

Following this introduction, Section 2 gives a brief overview of NDN architecture. Section 3 thoroughly discusses Interest flooding attack, and Section 4 proposes multiple defensive techniques against it. Section 5 describes evaluation methodology and gives experimental results. Section 6 contains related work. Section 7 provides analysis based on our results. Section 8 concludes and describes future work.

%\begin{figure}[htpb]
%  \centering
%  \includegraphics[scale=0.5]{figures/sim-emu-power.pdf}
%  \caption{Strength of Interest flooding attack}
%  \label{fig:simemupower}
%\end{figure}

%\begin{figure}[htpb]
%  \centering
%  \includegraphics[scale=0.5]{figures/sim-emu-performance.pdf}
%  \caption{Data retrieval by legitimate clients}
%  \label{fig:simemuperf}
%\end{figure}

>>>>

% Introduction to the problem: ndn, solves many problems, but has potential to introduce new problems.
% in this paer we are studing a number of remedies to Interest floodig attack, as well as trying to doscover ptential of tjese dolutions and their shortcomings

% Structure. Hopefully around three paragraps. 
% - what is happening with ndn and what is interest floodog
% - what are the general approaches we are taking, ranging from naïve to more intelligent
% - basic summary of the results

% move disclaimer from section 3 to the intro.

Various forms of distributed denial of services attack (DDoS) pose a significant and constant threat to the existing Internet infrastructure~\cite{arbor-report}.
The recently proposed Named Data Networking (NDN) architecture~\cite{ndn-conext, ndn-tr} completely changes the communication paradigm in the network by removing the notion of network host identities and making the application-level Data names the first-class network citizens.
That is, instead of requiring users to explicitly connect to specific servers and then use application-level protocols to request desired Data, users can directly ask the network for the Data.
% In other words, NDN changes communication from host-centric to data-centric.
Such a shift automatically solves several long standing problems, including source address spoofing and reflector DDoS attacks~\cite{mirkovic2004taxonomy}.

However, as long as there is still a notion of destination, requests (Interests in NDN terms) can still be send towards the destinations,%
\footnote{Note that ```destination'' is much more vague in NDN than in IP. 
Copies of the requested Data, towards which Interests may be forwarded, can be in many places, including long-term and short-term caches on NDN routers and alternative locations of the Data producer.} 
potentially disrupting service to legitimate users.
Taking into account that NDN requires the stateful forwarding of Interets---as the only way for routers to return Data back is to remember from where the Interest came in---an attacker may specifically target the stateful forwarding of NDN, e.g., in a form of Interest flooding attack.

Our primary goal in this paper is not to completely solve all denial of service problems in NDN, but to demonstrate using the Interest flooding attack example (Section~\ref{sec:interest flooding}) that although new NDN-specific DDoS attacks are possible in NDN network, the network is fully equipped to mitigate effects of these attacks.
In the paper we design several algorithms to mitigate the Interest flooding attack (Section~\ref{sec:design}), exploring the following two unique venues provided by NDN.
First, an NDN router does not need to accept, create state, and forward all the incoming Interests, but only the amount that would fully utilize downstream link capacity.
Second, the fact that Data always follows the Interest paths can be used to distinguish between good (e.g., a forwarded Interest gets satisfied) and ``bad'' Interests.

% Alex: Not sure whether it should be mentioned in intro or not.  If it should, then I don't really like it...
To quantify the quality of the designed attack mitigation algorithms, we performed a simulation-based study (Section~\ref{sec:evaluation}) based on trivial small-scale and Internet-like larger-scale topologies.
The results show a great potential of one of the designed algorithms (dynamic limits, Section~\ref{sec:dynamic limits}), where the attack traffic can be effectively localized near the attacker himself.

% Section 2 gives a brief overview of NDN architecture. 
% Section 3 defines Interest flooding attack, 
% Section 4 designes several attack mitigation algorithms.
% Section 5 describes evaluation methodology and gives experimental results. Section 6 contains related work. Section 7 provides analysis based on our results. Section 8 concludes and describes future work.

### 2012-12-09 03-interest-flooding.tex
Most of the evaluation section is semi-done.  I suspect I put too much details and discussion in the evalutions... Not sure
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@871 df1a90cd-247b-4364-bdaf-287a704eb499
New intro, updated discussion
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@876 df1a90cd-247b-4364-bdaf-287a704eb499

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\columnwidth]{attack-definition}
  \caption{Example of the Interest flooding attack}
  \label{fig:flooding example}
\end{figure}

@@

\paragraph{Disclaimer}

In this paper we are not trying to completely solve all denial of service problems that can be caused by various variations of the Interest flooding attack.
Our primary focus is to demonstrate that while NDN opens doors to new types of denial of service attacks, stateful packet forwarding of NDN makes it possible to mitigate such attacks, adding little additional complexity to the network.
In section~\ref{sec:discussion} we discuss our choice of the assumptions in more detail, as well as discuss potential of the Interest flooding attack under several other attack assumptions.

>>>>

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.39]{attack-definition}
  \caption{Example of the Interest flooding attack}
  \label{fig:flooding example}
\end{figure}

@@

In section~\ref{sec:discussion} we discuss potential of the Interest flooding attack under several other attack assumptions.

### 2012-12-09 07-discussion.tex
New intro, updated discussion
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@876 df1a90cd-247b-4364-bdaf-287a704eb499

% Lessons we learnt, comparison of different mitigation techniques. 
% Adapting these to mitigate other DDoS attacks.

% In section~\ref{sec:interest flooding} we made assumptions regarding the Interest flooding traffic shape (requesting non-existing data with non-aggregatable prefix name ) and single-path forwarding strategy, which jointly create the most powerful congestion at the producer link.

% In this paper we are using a simple attacker profile and greedy attacker, which is using all resources (available bandwidth) of its botnet in order to put producer down as quickly as possible. However, in a real life not all botnets are greedy and in fact can perform smart actions like constant switching between victims every other hour to fool defensive systems by nullifying gathered bad flow statistics. Or they can be non-greedy at all, for instance a large botnet of several hundreds of thousands nodes can perform a non-greedy attack by sending just a few Interests per second from each node. Nowadays, this situation is very hard to manage, because of the difficulties in detecting of bad flows in a traffic.

% Also, an adversary can mix Interests for non-existing data with Interests for existing-data. Thanks to NDN stateful packet forwarding, bad flows can be easily distinguished and penalized.

@@

The attack mitigation methods studied in this paper provide ways to eliminate network congestion and router overload problems caused by junk Interest flooding (Section~\ref{sec:interest flooding}), but unfortunately not without inducing some new problems.
In particular, during an ongoing attack, a number of legitimate Interests successfully forwarded towards the Data producer can be substantially larger in unprotected network, then in the network that implements Interests mitigation algorithms.  
This is simply because an unprotected network allows clients to pump more packets and does not impose any artificial restrictions on packet forwarding.
Therefore, unless a mitigation algorithm uses some form of intelligence (e.g., statistics for Interest satisfiability), a smaller scale flooding attack can cause greater damage to the service quality (recall service quality curves for physical limits in small-scale and larger-scale evaluations, Fig.~\ref{fig:small-scale} and Fig.~\ref{fig:large-scale}).

Let us take one small step back.
In a general case, the adversaries would not be aware that the network is or is not protected, e.g., using dynamic the limits algorithm.
As a result, they will try to do as much harm as possible using junk Interests and the mitigation algorithm will work as expected: small service quality degradation during first seconds of the attack, followed by practically full recovery from the attack.
However, a careful reader may ask a totally legitimate question: what would happen if adversaries know that the network is protected and know that it is using the dynamic limits algorithm?
Would there be an optimal strategy to deny service for the legitimate users?
Would it be better to send ``legitimate'' Interests instead, at least partially, and attempt to always keep the network in the transitional state?

% Hopefully, I'll have some results to back up my claims, otherwise these are just words

% We acknowledge that this could be a probl
Although a complete investigation of this problem deserves a separate study, we anticipate that the problem may not be as severe as it seems.
When attacker's Interests can be satisfied (e.g., an attacker requesting dynamically generated Data), these Interest would not linger on routers for extended periods of time, giving opportunities for Interests from the legitimate users to get through and bring Data back.
Essentially, in the worst case scenario (assuming that dynamic limits are augmented with some form fair queuing), legitimate users and attackers would ``fairly'' share the network.
(Note that router resource overloading is still impossible, as there is always an upper bound of admitted Interets.)

>>>>

@@

The attack mitigation methods studied in this paper provide ways to eliminate network congestion and router overload problems caused by junk Interest flooding (Section~\ref{sec:interest flooding}), but unfortunately not without inducing some new problems.
In particular, during an ongoing attack, a number of legitimate Interests successfully forwarded towards the Data producer can be substantially larger in unprotected network, then in the network that implements an Interest mitigation algorithm.  
This is simply because an unprotected network allows clients to pump more packets and does not impose any artificial restrictions on packet forwarding.
Therefore, unless a mitigation algorithm uses some form of intelligence (e.g., statistics for Interest satisfiability), a smaller scale flooding attack can cause greater damage to the service quality (recall service quality curves for physical limits in small-scale and larger-scale evaluations, Fig.~\ref{fig:small-scale} and Fig.~\ref{fig:large-scale}).

Although the studied attack mitigation algorithms base their intelligence on per interface level, it is also possible to use more granular processing.
For example, if routers measure satisfaction ratios and limit Interest forwarding independently for each FIB prefix, then an ongoing attack on one Data namespace will not have any effect ob users requesting Data from another namespace.
 % (assuming that users have to get through te same roiter to get the data)

% (already mentioned in Interest flooding definiton section)
%  Another thing is that rich interconnectivity (de facto on The Internet) and the multi-path packet forwarding (practically impossible) gives NDN addition leverage points to mitigate / reduce effects of ongoing Interest flooding attack of any form.

In the definition of the Interests flooding attack we explicitly stated that an attack is carried using junk Interets, which obviously gives us a clear distinction between ``good'' and ``bad'' Interests.
A curious reader may ask us a valid question: what would happen if an adversary doing something else, e.g., sending sends only ``legitimate'' Interests or intermix them with junk Interests?
Although a complete investigation of this problem deserves a separate study, we anticipate that the problem may not be as severe as it seems.
First, in a general case, the adversaries would not be aware that the network is or is not protected, e.g., is using the dynamic limits algorithm.
As a result, adversaries would want to do as much harm as possible, which is to send junk Interests.

% Second, when adversary is aware of the attack mitigation mechanisms and sends 
Second, a satisfiable Interests does not cause NDN-specific harm, as they would not linger on routers for extended periods of time, giving opportunities for Interests from the legitimate users to get through and bring Data back.
Essentially, in the worst case scenario (assuming that dynamic limits are augmented with some form fair queuing), legitimate users and attackers would ``fairly'' share the network.
Note that router resource, as well as upstream link capacity overloading is still impossible, as there is always an upper bound of admitted Interets.

%  and the mitigation algorithm will work as expected: small service quality degradation during first seconds of the attack, followed by practically full recovery from the attack.
% However, a careful reader may ask a totally legitimate question: what would happen if adversaries know that the network is protected and know that it is using the dynamic limits algorithm?
% Would there be an optimal strategy to deny service for the legitimate users?
% Would it be better to send ``legitimate'' Interests instead, at least partially, and attempt to always keep the network in the transitional state?
